Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2016-12-20 manuscript/reviewer_response_tree_physiology.Rmd

We are grateful for the opportunity to resubmit our manuscript. We thank the reviewers for providing insightful comments to improve our manuscript. We have responded in detail to each comment, which has improved the overall quality of the manuscript. Below, we have itemized our response to each comment from both referees.

### **Referee #1**

\
- *Reviewer: 1. The role of nutrient defiency may have been underestimated...The relative contribution should be determined...*

\
*Response*: Unfortunately, we do not have the ability with these data to determine the relative contribution of nutrient deficiency on the down regulation of photosynthesis. We were able to identify not one, but two processes that likely contributed to photosynthetic inhibition.  Both of these processes, nutrient limitation and starch accumulation, are discussed at length in the discussion. We absolutely agree that determining the contribution of each of these processes will provide valuable insights on how sink limitation impacts the carbon balance of plants. By providing convincing empirical evidence of these processes, this experimental design should provide a framework for future work to quantify the contributions of each.

@@

\
- *Reviewer: 6. pp 63: “…the sink limitation of growth has been explored by manipulations of C source and sink activity…. Examples of …that manipulate C source activity include… and partial defoliation..” (partial defoliation should impose a sink limitation to growth?) Compensatory increases of An are commonly found, but not allowing a higher C net gain leading to sink limitation.*

\
*Response*: Regardless of the downstream effects of defoliation on growth processes, the removal of leaves is first and foremost a source manipulation. This has been repeatedly shown and is a commonly utilized manipulation in fruit trees and crops species.

\
- *Reviewer: 7. pp 365: No changes in stomatal conductance during the experiment in the described field conditions (Figure 2)?*

\
*Response*: Soil moisture for each seedling was maintained at field capacity, throughout the duration of the experiment, in order to minimize affects of reduced water availability on photosynthesis and growth. The consistent lack of water stress on theses seedlings (see Discussion "Changes in growth and physiology under sink limitation") was an successful implement of the experimental design.

@@

\
- *Reviewer: 9. Table 1 and 2: Reorganize. Biomass, SLA, SRL and FRLD in one table, and chemistry parameters in a second table.*

\
*Response*: We agree with the reviewer. Table 1 and 2 have been reorganized, as suggested, to match how the data are presented in the Results section.

\
- *Reviewer: 10. Figure 4: Significant differences between 5 and 15 L soil volumes? Are SE or SD displayed?*

\
*Response*: Bar plots of Figure 4 are means and standard errors which are described in figure caption. Yes, the 5 L containers had smaller rates of A~sat~ than either 15 or 20 L containers. The 5 L containers also had the lowest measured rates overall, indicative of a strong impact of a small container size on sink strength.  For simplicity, we report the large overall difference in Free seedlings and seedlings in containers in terms of photosynthesis rates.

@@

\
- *Reviewer: 4. In “Leaf, root and soil chemistry” indicate how plant samples were dried (temperature, drying time), how they were milled (type of mill used, model, etc.) and how soils were sieved (pore size) and milled (type of mill used, model, etc.)*

\
*Response*: We have clarified in “Leaf, root and soil chemistry” that samples were over dried at 60&deg;C to a constant mass and ball milled for analysis. We have also added the pore size, drying time and mill type to the explanation of soil processing.

\
- *Reviewer: 5. Clairify how did soil volume treatments relate to the 61 digitalized seedlings, which include different Eucalyptus species? Were the 61 digitized seedlings grown in different soil volumes*

\
*Response*: The digitized seedlings used here are a collection of digitized plants from differing environments including understory, glasshouse and plantation (see Duursma et al. 2012).  We selected 5 *Eucalpytus* species from this collection, which represent canopy structures of similar seedling types and sizes. Although the digitized plants are from separate experiments, they allowed us to estimate self shading of the *Eucalyptus* seedlings in this experiment from regressions with total plant leaf area and canopy dimensions (cf. Duursma et al. 2012). As this experiment had a ~4 month duration, leaf counts and total leaf area of seedlings were generally small. This resulted in only a modest adjustment of the self-shading factor to total canopy carbon uptake.

### 2017-01-12 manuscript/manuscript_pve.Rmd
quick html file to Remko to finalize A-tnc-N reviewer response. Will integrate these edits into response letter and Results section

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the Free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ (Figure `r asatfig`) and *A*~max~ (Table `r pvetab3`) were consistently higher in Free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was highly correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~max~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~max~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~max~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

@@

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although *g*~s~ in Free seedlings was generally higher than those in containers (Table `r pvetab3`, *P* < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table `r pvetab2`). Combined, these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N at harvest was not different across soil volumes (mean of .045 %), with minimal decreases from pre-planting values (mean of = .049 %). This similarity indicates that nutrient leaching from Free seedlings or from draining of containers following natural rainfall events did not differ between treatments. 
##Whole-plant C balance
Modelled cumulative net leaf C gain over 120 days varied three-fold across containers sizes (Table `r pvetab1`). The Free seedling control had five-fold greater modelled total C gain than the average of container treatments (213.6 - 42.7&pm;8.2 g C). For seedlings in containers, an average of 67 &pm;0.01 % of modelled C gain was not allocated to observed biomass, compared to 59.2 % for the Free seedling control. Consequently, the estimated CUE of containerized seedlings was lower than the Free seedling control (Figure `r cuefig`). As a result, the observed reductions in leaf *A*~n~ with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the container size treatments.

>>>>

##Gas exchange and photosynthetic parameters
At the first measurement campaign, both *A*~sat~ and *A*~max~ were significantly higher in the Free seedling treatment compared to seedlings in containers (both *P* < 0.001). Across all measurement campaigns, mean *A*~sat~ (Figure `r asatfig`) and *A*~max~ (Table `r pvetab3`) were consistently higher in Free seedlings than in containers (26 % and 29 %, respectively). The relationship between leaf starch and leaf N~f~ with *A*~max~ on a mass basis was marginally significant (*P* = 0.058), but *A*~max~ on a mass basis was highly correlated with both leaf N~f~ and leaf starch (both *P* < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationships of *A*~max~ to either leaf N~f~ or leaf starch at multiple bin levels (n=5) of the co-variate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments *A*~max~ was higher when leaf N~f~ was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). *A*~max~  was also lower when leaf starch was high as higher leaf N~f~ often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, *A*~max~ was positively correlated with final harvest biomass across all seedlings (*P* < 0.001).

@@

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although *g*~s~ in Free seedlings was generally higher than those in containers (Table `r pvetab3`, *P* < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table `r pvetab2`). Combined, these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N at harvest was not different across soil volumes (mean of .045 %), with minimal decreases from pre-planting values (mean of = .049 %). This similarity indicates that nutrient leaching from Free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

##Whole-plant C balance
Modelled cumulative net leaf C gain over 120 days varied three-fold across containers sizes (Table `r pvetab1`). The Free seedling control had five-fold greater modelled total C gain than the average of container treatments (213.6 - 42.7&pm;8.2 g C). For seedlings in containers, an average of 67 &pm;0.01 % of modelled C gain was not allocated to observed biomass, compared to 59.2 % for the Free seedling control. Consequently, the estimated CUE of containerized seedlings was lower than the Free seedling control (Figure `r cuefig`). As a result, the observed reductions in leaf *A*~n~ with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the container size treatments.
